---
title: "bundestag_project"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(stargazer)
```


```{r}
rawdata <-read.csv("bundestag.csv",sep=";")

head(rawdata)

# delete first,5,6 colum
data_col_clean <- subset(rawdata,select = -c(1,5,6))

head(data_col_clean)

deutschland_sum <-data_col_clean[1,]
deutschland_sum
# put deutschland extra 
data_row_clean <- data_col_clean[-c(1),]
data_row_clean
# select all rows which have a two digit first column (bundesländer) (+delete these rows)
data_row_clean$X.1 <- as.integer(data_row_clean$X.1) 
data_no_bundesland <-subset(data_row_clean, X.1 == 11 | X.1==2| X.1< 20000 &  X.1 > 999)
data_no_bundesland <-subset(data_no_bundesland ,X.3!="-")
data_no_bundesland

data_no_bundesland$X.3<- as.integer(data_no_bundesland$X.3)
wahlberechtige_sum  <- sum(data_no_bundesland$X.3)
wahlberechtige_sum
data_no_bundesland <-subset(data_no_bundesland ,select=-c(3))
data <-subset(data_no_bundesland ,select=-c(1))

colnames(data) <- c("X.2","CDU/CSU","SPD","GRÜNE","FDP","DIE LINKE","AFD","OTHER")
data
```



```{r}
## change to right category 
data[, 2:8] <- sapply(data[, 2:8], as.integer)
data
wahl <- subset(data ,select=-c(1))
wahl
sums <- rowSums(wahl)
wahlsums <- cbind(wahl,  sums)
wahlsums
dim(wahlsums)
wahl_percentage <-  sweep(wahl,1,sums,'/')
wahl_percentage


wahl2 <- cbind(wahl_percentage,  rowSums(wahl_percentage))
wahl2

```
```{r}
# get bundesland for each 
bundesland <-subset(data_no_bundesland ,select=c(1))
bundesland
bundesländer <- c("Schleswig-Holstein","Hamburg","Niedersachsen","Bremen","Nordrhein-Westfalen","Hessen","Rheinland-Pfalz","Baden-Württemberg","Bayern","Saarland","Berlin","Brandenburg","Mecklenburg-Vorpommern","Sachsen","Sachsen-Anhalt","Thüringen")
werte <- c(1000,2000,3000,4000,5000,6000,7000,8000,9000,10000,11000,12000,13000,14000,15000,16000)
bundesländer
# 
#whichbundesland <- vector(mode="integer", length=len(data_no_bundesland$X.1))
who <- trunc((data_no_bundesland$X.1/1000))
who[16]<- 2
who[325]<- 11
who
neu <- vector(mode="integer", length=length(who))
# neue und alte bundesländer Brandenburg, Mecklenburg-Vorpommern, Sachsen, Sachsen-Anhalt und Thüringen
for(i in 1:length(who)){
  if(who[i] >11){
    neu[i] = 2
  }else{
    neu[i] = 1
  }
}
```

```{r}
sum(is.na(wahl_percentage))
```
```{r}
summary(wahl_percentage)

```

## Single variate analysis
# SAARLAND BEEING DUMB ALSO VISUALIZE 

```{r}

means = colMeans(wahl_percentage)
medians =apply(wahl_percentage, 2, FUN = median)
standards =  apply(wahl_percentage,2,sd)


means
medians
standards

```
```{r}
library(scales)
```


```{r}
pairs(wahl_percentage, pch = c(3:18)[who], gap = 0, upper.panel = NULL,col=alpha(rainbow(16)[who], 0.4))
par(xpd = TRUE)
legend(0.8,0.95,legend = bundesländer,pch =c(3:18),col=rainbow(16),cex = 0.5)

```

```{r}

pairs(wahl_percentage, pch = c(3:18)[who], gap = 0, upper.panel = NULL,col= alpha(rainbow(2)[neu], 0.4))
par(xpd = TRUE)
legend(0.8,0.95,legend = c("alt","neu"),fill =rainbow(2),cex = 0.5)

```
```{r}

length(wahl_percentage)

```

## maybe add some weighting or smth 


```{r}
wahl_dist<- dist(wahl_percentage,method="euclidian",diag=TRUE,upper=FALSE)
```


```{r}
par(mfrow = c(1, 3))
plot(hclust(wahl_dist, method = "single"), main = "Single linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
plot(hclust(wahl_dist, method = "average"), main = "Average linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
plot(hclust(wahl_dist, method = "complete"), main = "Complete linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
```
```{r}

library(dendextend)
```
```{r}
clustering <- hclust(wahl_dist, method = "average")
dend <- as.dendrogram(clustering)
labels_colors(dend) <- rainbow(16)[who][order.dendrogram(dend)]
plot(dend)
```
---
title: "bundestag_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
;;;Wahlberechtigte;Wahlbeteiligung;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen;Gültige Zweitstimmen
;;;;;;Parteien;Parteien;Parteien;Parteien;Parteien;Parteien;Parteien
```{r}
rawdata <-read.csv("bundestag.csv",sep=";")

head(rawdata)

# delete first,5,6 colum
data_col_clean <- subset(rawdata,select = -c(1,5,6))

head(data_col_clean)

deutschland_sum <-data_col_clean[1,]
deutschland_sum
# put deutschland extra 
data_row_clean <- data_col_clean[-c(1),]
data_row_clean
# select all rows which have a two digit first column (bundesländer) (+delete these rows)
data_row_clean$X.1 <- as.integer(data_row_clean$X.1) 
data_no_bundesland <-subset(data_row_clean, X.1 == 11 | X.1==2| X.1< 20000 &  X.1 > 999)
data_no_bundesland <-subset(data_no_bundesland ,X.3!="-")
data_no_bundesland

data_no_bundesland$X.3<- as.integer(data_no_bundesland$X.3)
wahlberechtige_sum  <- sum(data_no_bundesland$X.3)
wahlberechtige_sum
data_no_bundesland <-subset(data_no_bundesland ,select=-c(3))
data <-subset(data_no_bundesland ,select=-c(1))
data
#column
```



```{r}
## change to right category 
data[, 2:8] <- sapply(data[, 2:8], as.integer)
data
wahl <- subset(data ,select=-c(1))
wahl
sums <- rowSums(wahl)
wahlsums <- cbind(wahl,  sums)
wahlsums
dim(wahlsums)
wahl_percentage <-  sweep(wahl,1,sums,'/')
wahl_percentage


wahl2 <- cbind(wahl_percentage,  rowSums(wahl_percentage))
wahl2

```
```{r}
# get bundesland for each 
bundesland <-subset(data_no_bundesland ,select=c(1))
bundesland
bundesländer <- c("Schleswig-Holstein","Hamburg","Niedersachsen","Bremen","Nordrhein-Westfalen","Hessen","Rheinland-Pfalz","Baden-Württemberg","Bayern","Saarland","Berlin","Brandenburg","Mecklenburg-Vorpommern","Sachsen","Sachsen-Anhalt","Thüringen")
werte <- c(1,2,3,4,5,6,7,8,9,1,11,12,13,14,15,16)
bundesländer
# 
#whichbundesland <- vector(mode="integer", length=len(data_no_bundesland$X.1))
who <- trunc((data_no_bundesland$X.1/1000))
who[16]<- 2
who[325]<- 11
who
neu <- vector(mode="integer", length=length(who))
# neue und alte bundesländer Brandenburg, Mecklenburg-Vorpommern, Sachsen, Sachsen-Anhalt und Thüringen
for(i in 1:length(who)){
  if(who[i] >11){
    neu[i] = 2
  }else{
    neu[i] = 1
  }
}
```

```{r}
sum(is.na(wahl_percentage))
```
```{r}
summary(wahl_percentage)

```


```{r}
boxplot(wahl_percentage)
```

## Single variate analysis
# SAARLAND BEEING DUMB ALSO VISUALIZE 

```{r}

means = colMeans(wahl_percentage)
medians =apply(wahl_percentage, 2, FUN = median)
standards =  apply(wahl_percentage,2,sd)


means
medians
standards

```
```{r}
library(scales)
```


```{r}
pairs(wahl_percentage, pch = c(3:18)[who], gap = 0, upper.panel = NULL,col=alpha(rainbow(16)[who],0.4))
par(xpd = TRUE)
legend(0.85,0.95,legend = bundesländer,pch =c(3:18),col=rainbow(16),cex = 0.5)
```


```{r}
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)


legend(0.5,0.5,legend = bundesländer,pch =c(3:18),col=rainbow(16),cex = 1.0)

```

```{r}

pairs(wahl_percentage, pch = c(3:18)[who], gap = 0, upper.panel = NULL,col= rainbow(2)[neu])
par(xpd = TRUE)
legend(0.8,0.95,legend = c("alt","neu"),fill =rainbow(2),cex = 0.5)

```
```{r}

length(wahl_percentage)

```

## maybe add some weighting or smth 


```{r}
wahl_dist<- dist(wahl_percentage,method="euclidian",diag=TRUE,upper=FALSE)
```


```{r}
par(mfrow = c(1, 3))
plot(hclust(wahl_dist, method = "single"), main = "Single linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
plot(hclust(wahl_dist, method = "average"), main = "Average linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
plot(hclust(wahl_dist, method = "complete"), main = "Complete linkage",
xlab = "Region / Clusters", sub = NA, hang = -1)
```


```{r}

library(dendextend)
```



```{r}
clustering <- hclust(wahl_dist, method = "average")
dend <- as.dendrogram(clustering)
labels_colors(dend) <- rainbow(16)[who][order.dendrogram(dend)]
plot(dend)
```


```{r}
clustering2 <- hclust(wahl_dist, method = "average")
dend <- as.dendrogram(clustering2)
labels_colors(dend) <- rainbow(2)[neu][order.dendrogram(dend)]
plot(dend)
```


```{r}
plot(dend,horiz = TRUE)
```
```{r}
clusters <- cutree(clustering, h = 0.15)
clu <- data.frame(clusters,who)
clu
cluwho <- data.frame(data$X.2,clusters,who)
cluwho

cluall <- data.frame(clusters,who,data)
cluall

tabdf1 <- as.data.frame.matrix(t(table(clu)))
tabdf1
rownames(tabdf1)<- bundesländer
tabdf1
```
```{r}
library(gridExtra)
```

```{r}
stargazer(tabdf1, type = "latex")
```



```{r}

clusters2 <- cutree(clustering2, h = 0.22)
clu2 <- data.frame(clusters2,who)
table(clu2)
table(clusters2)
tabdf <- as.data.frame.matrix(t(table(clu2)))
rownames(tabdf)<- bundesländer
tabdf
```




```{r}
stargazer(tabdf, type = "latex")
```



```{r}
cluall2 <- data.frame(clusters2,neu,data)
cluall2
```











